---
title: "04_describe"
format: 
  html:
    self-contained: true
editor: visual
---

## Load Libraries

```{r}
#| message: false
library("tidyverse")
library("here")
library("RColorBrewer")
source(here("R/99_proj_func.R"))
```

## Load Data

Load the data as a variable to use in descriptive graphs

```{r}
#| message: false
augment_data <- 
  read_tsv(file = here("data/03_dat_aug.tsv"))
```

## Descriptive Graphs

```{r}
custom_palette <- c(
  "#7f3667",
  # "#a53e76",
  "#bb437e",
  # "#d24787",
  "#e44b8d",
  "#e2619f",
  # "#e27bb1",
  "#e5a0c6",
  "lightblue",
  "#B3D84C",  
  # "#A6D15A",  
  "#90C84C",  
  "#70B83F",  
  "#4F9A2F",  
  # "#2F7B1F",  
  "#1D6510",  
  "#004400")
```

### Distribution of age groups and race

```{r}
df_plot1 <- augment_data |>
  mutate(
    age_group = cut(
      age,
      breaks = seq(0,
                   100,
                   by = 10),
      labels = paste(seq(0, 90, 10),
                     seq(9, 99, 10),
                     sep = "-"))) |>
  filter(
    !is.na(age_group)) |>
  count(age_group,
        race)

plot1_pallete <- c("#e44b8d",
                   "#e5a0c6",
                   "lightblue",
                   "#B3D84C",
                   "#70B83F")

gg_4_1 <- ggplot(
  df_plot1,
  aes(
    x = age_group,
    y = n,
    fill = race)) +
  geom_col(position = "stack") +
  labs(
    title = "Distribution of Age Groups and Race",
    x = "Age groups",
    y = "Number of subjects",
    fill = "Races") +
  scale_fill_manual(values = plot1_pallete) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

gg_4_1
```

### Distribution of hormones in different patients

```{r}
df_plot2 <- augment_data |>
  mutate(
    receptor_group = case_when(
      er_status == "N" & pr_status == "N" & her2_status == "N" ~ "ER-PR-HER2-",
      er_status == "P" & pr_status == "N" & her2_status == "N" ~ "ER+PR-HER2-",
      er_status == "N" & pr_status == "P" & her2_status == "N" ~ "ER-PR+HER2-",
      er_status == "N" & pr_status == "N" & her2_status == "P" ~ "ER-PR-HER2+",
      er_status == "P" & pr_status == "P" & her2_status == "N" ~ "ER+PR+HER2-",
      er_status == "P" & pr_status == "N" & her2_status == "P" ~ "ER+PR-HER2+",
      er_status == "N" & pr_status == "P" & her2_status == "P" ~ "ER-PR-HER2-",
      er_status == "P" & pr_status == "P" & her2_status == "P" ~ "ER+PR+HER2+")) |>
  count(receptor_group)

plot2_palette <- c("#7f3667",
                   "#e44b8d",
                   "#e5a0c6",
                   "lightblue",
                   "#B3D84C", 
                   "#70B83F",
                   "#2F7B1F")

gg_4_2 <- ggplot(
  df_plot2,
  aes(
    x = receptor_group,
    y = n,
    fill = receptor_group)) +
  geom_col() +
  labs(
    title = "Distribution of Hormones",
    x = "Receptor groups",
    y = "Number of occurances",
    fill = "Receptor groups") +
  scale_fill_manual(values = plot2_palette) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 30,
                                   hjust = 0.9,
                                   vjust = 1.2),
        plot.title = element_text(hjust = 0.5))

gg_4_2
```

### Distribution of successful or unsuccessful treatment

```{r}
df_plot3 <- augment_data |>
  filter(
    !is.na(response)) |>
  count(response)

plot3_palette <- c("#e44b8d",
                   "#90C84C")

gg_4_3 <- ggplot(
  df_plot3,
  aes(
    x = response, 
    y = n,
    fill = response)) +
  geom_col() +
  labs(
    title = "Distribution of Treatment Success: 
             pCR = 'no detectable tumor after treatment'
             RD = 'tumor remains after therapy'",
    x = "Response type", 
    y = "Number of cases",
    fill = "Response type") +
  scale_fill_manual(values = plot3_palette) +
  theme_minimal()

gg_4_3
```

### Distribution of tumor subtypes (histology)

```{r}
df_plot4 <- augment_data |>
  filter(
    !is.na(histology)) |>
  count(histology)

gg_4_4 <- ggplot(
  df_plot4,
  aes(
    x = histology, 
    y = n,
    fill = histology)) +
  geom_col() +
  labs(
    title = "Distribution of Tumor Subtypes (Histology)",
    x = "Histology",
    y = "Number of cases",
    fill = "Histology") +
  scale_fill_manual(values = custom_palette) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1,
                                   vjust = 1),
        plot.title = element_text(hjust = 0.5))

gg_4_4
```

### Distribution of tumor aggressiveness (bmngrd = Bloom–Richardson (BR) grade)

```{r}
df_plot5 <- augment_data |>
  mutate(grade = factor(grade)) |>
  count(grade)

plot5_palette <- c("#e44b8d",
                   "lightblue",
                   "#90C84C")

gg_4_5 <- ggplot(
  df_plot5,
  aes(
    x = grade, 
    y = n,
    fill = grade)) +
  geom_col() +
  labs(
    title = "Distribution of Tumor Aggressiveness (BR grade)",
    x = "Grade",
    y = "Number of cases", 
    fill = "Grade") +
  scale_fill_manual(values = plot5_palette, na.value = "gray") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))

gg_4_5
```

### Save plots as png

```{r}
save_plot(gg_4_1, "04_key_plot_1.png")
save_plot(gg_4_2, "04_plot_2.png")
save_plot(gg_4_3, "04_plot_3.png")
save_plot(gg_4_4, "04_plot_4.png")
save_plot(gg_4_5, "04_plot_5.png")
```
