---
title: "19_analysis_15"
format: html
editor: visual
---

## Running Code

#### Loading libraries:

```{r}

library("dplyr")  
library("ggplot2")  
library("scales")
```

#### Read data:

```{r}

dat <- read.delim("../data/03_dat_aug.tsv")
```

#### Re-ordering

Ordering the different genotypes into groups from the augmented data:

```{r}

# Create ER/PR/HER2 combo groups
dat <- dat |>
  mutate(
    ER   = if_else(er_status  == "P",
                   "ER+",
                   "ER-"),
    PR   = if_else(pr_status  == "P", 
                   "PR+", 
                   "PR-"),
    HER2 = if_else(her2_status == "P",
                   "HER2+",
                   "HER2-"),
    combo = paste(ER, 
                  PR,
                  HER2,
                  sep = ", ")
  ) |>
  mutate(
    combo = factor(
      combo,
      levels = c(
        "ER+, PR-, HER2-",
        "ER+, PR+, HER2-",
        "ER+, PR-, HER2+",
        "ER+, PR+, HER2+",
        "ER-, PR+, HER2-",
        "ER-, PR+, HER2+",
        "ER-, PR-, HER2+",
        "ER-, PR-, HER2-"
      )
    )
  )

# check: how many rows total?
nrow(dat)

# check: how many patients in each genotype?
dat |> count(combo)
```

#### **Graph 2:**

##### Tumor Grade (1–3) by ER/PR/HER2 Genotypes

```{r}

# Prepare data with grade as factor (1,2,3,NA)
dat_grade <- dat |>
  mutate(
    grade = as.character(grade),
    grade = if_else(is.na(grade),
                    "NA",
                    grade),
    grade = factor(grade,
                   levels = c("1", "2", "3", "NA"))
  )

# total N (same as before, but explicit)
N_total_grade <- nrow(dat_grade)

# N per genotype
group_counts_grade <- dat_grade |>
  group_by(combo) |>
  summarise(N_group = n(),
            .groups = "drop")

ggplot(dat_grade,
       aes(
         x = combo,
         fill = grade)) +
  geom_bar(position = "fill") +
  scale_x_discrete(drop = FALSE) +
  
  # Red colour scheme + grey for NA
  scale_fill_manual(
    values = c(
      "1" = "#FFCCCC",  # very light red
      "2" = "#FF6666",  # medium red
      "3" = "#CC0000",  # dark red
      "NA" = "grey80"   # NA grey
    ),
    drop = FALSE,
    name = "Grade"
  ) +
  scale_y_continuous(labels = percent_format()) +
  
  # N labels above bars
  geom_text(
    data = group_counts_grade,
    aes(
      x = combo,
      y = 1.05,
      label = paste0("N=", N_group)
    ),
    size = 3,
    inherit.aes = FALSE
  ) +
  labs(
    title = "Tumor Grade (1–3) by ER/PR/HER2 Genotypes",
    x = "ER/PR/HER2 Genotypes",
    y = paste0("Proportion of Patients (N = ", N_total_grade, ")"),
    fill = "Grade"
  ) +
  theme_minimal() +
  theme(
    axis.text.x  = element_text(angle = 45,
                                hjust = 1,
                                size = 8),
    plot.title   = element_text(size = 10),
    axis.title.y = element_text(size = 8),
    axis.title.x = element_text(size = 8)
  ) +
  coord_cartesian(ylim = c(0, 1.15))
```

##### **Conclusions:**

Even tough being postitive for Estrogen receptors do have the worst outcome (Graph 1), they do not have a more agressive tumor grade (Graph 2). Being negative for both hormone receptors leads to a more aggressive tumor grade (Graph 2), and those two actually had the worst outcome (Graph 1).
