---
title: "20_analysis_16"
format: html
editor: visual
---

## Running Code

#### Loading libraries:

```{r}
#| message: false
library("here")
library("tidyverse")
library("scales")
```

#### Read data:

```{r}
#| message: false
dataset <- read_tsv(file = here("data/03_dat_aug.tsv"))
```

#### Re-ordering

Ordering the different genotypes into groups from the augmented data:

```{r}

# Create ER/PR/HER2 combo groups
dataset <- dataset |>
  mutate(
    ER = if_else(
      er_status == "P", 
      "ER+",
      "ER-"),
    PR = if_else(
      pr_status == "P",
      "PR+", 
      "PR-"),
    HER2 = if_else(
      her2_status == "P",
      "HER2+",
      "HER2-"),
    combo = paste(
      ER,
      PR,
      HER2,
      sep = ", ")) |>
  mutate(
    combo = factor(
      combo,
      levels = c(
        "ER+, PR-, HER2-",
        "ER+, PR+, HER2-",
        "ER+, PR-, HER2+",
        "ER+, PR+, HER2+",
        "ER-, PR+, HER2-",
        "ER-, PR+, HER2+",
        "ER-, PR-, HER2+",
        "ER-, PR-, HER2-")))

# check: how many rows total?
nrow(dataset)

# check: how many patients in each genotype?
dataset |> 
  count(combo)
```

#### Graph 3:

##### **Prepare for heatmap plot to look at the tumor size (0-4) among genotypes:**

```{r}

# 1. Make a clean tumor-size variable with NA as category
dataset_ts <- dataset |>
  mutate(
    ts = ifelse(
      is.na(tumor_size_before),
      "NA",
      as.character(tumor_size_before)))

# all genotype levels (from your combo factor) and all tumor-size levels
all_combos <- levels(dataset$combo)
ts_levels  <- c("0", "1", "2", "3", "4", "NA")    

# 2. Create full grid: every combo × every tumor-size level
full_grid_ts <- 
  expand.grid(
    combo = all_combos,
    ts = ts_levels,
    stringsAsFactors = FALSE)

# 3. Summary: counts & percentages per combo × tumor size
ts_summary <- 
  full_grid_ts |>
  left_join(
    dataset_ts |> 
      count(combo, ts),
    by = c("combo", "ts")) |>
  mutate(
    n = ifelse(
      is.na(n),
      0,
      n)) |>
  group_by(combo) |>
  mutate(
    total = sum(n),
    pct = ifelse(
      total > 0,
      n / total,
      0)) |>
  ungroup()

# factor for tumor size (plot rows)
ts_summary$ts <- factor(
  ts_summary$ts,
  levels = ts_levels)

# 4. N per combo (for labels) 
group_counts_ts <- 
  ts_summary |>
  distinct(combo,
           total) |>
  rename(N_group = total)

# 5. Extended factor for plotting: add an extra row "N" under 0 just for labels
ts_levels_with_N <- c(ts_levels, "N")  

ts_summary <- ts_summary |>
  mutate(
    ts_plot = factor(
      as.character(ts),
      levels = ts_levels_with_N))

n_label_df <- group_counts_ts |>
  mutate(
    ts_plot = factor(
      "N",
      levels = ts_levels_with_N))
```

##### **Graph - Heatmap plot -** Tumor Size Before Treatment by ER/PR/HER2 Genotypes

```{r}
gg_16 <- ggplot() +
  # heatmap tiles
  geom_tile(
    data = ts_summary,
    aes(
      x = combo,
      y = ts_plot,
      fill = pct)) +

  # percentage labels inside tiles
  geom_text(
    data = ts_summary,
    aes(
      x = combo,
      y = ts_plot,
      label = paste0(round(pct * 100),
                     "%")),
    size = 3) +

  # N labels in their own row under the heatmap
  geom_text(
    data = n_label_df,
    aes(
      x = combo,
      y = ts_plot,
      label = paste0("N=", 
                     N_group)),
    size = 3) +
  
  scale_fill_gradient(
    low   = "#e5a0c6",
    high  = "#4F9A2F",
    labels = percent_format(accuracy = 1),
    name  = "Proportion") +

  # keep all genotypes, even if some have few patients
  scale_x_discrete(drop = FALSE) +

  # y labels: hide label for the extra "N" row with ""
  scale_y_discrete(
    breaks = ts_levels_with_N,
    labels = c("0", "1", "2", "3", "4", "NA", "")) +

  labs(
    title = "Tumor Size Before Treatment by ER/PR/HER2 Genotypes",
    x     = "ER/PR/HER2 Genotypes",
    y     = "Tumor size before surgery") +
  theme_minimal() +
  theme(axis.text.x  = element_text(angle = 45, 
                                    hjust = 1, 
                                    size = 8),
        plot.title   = element_text(hjust = 0.5),
        axis.title.x = element_text(vjust = 0))

gg_16
```

```{r}
ggsave(gg_16, 
       filename = "20_key_plot_16.png",
       path = here("./results"),
       device = "png",
       height = 5,
       width = 8)
```

##### Conclusion:

Far most of the genotype groups have a size 2 tumor stage, or otherwise they have developed into size 3-4. This is especially the case for groups that are Estrogen receptor positive, and the last two groups which are negative for both hormone receptors.
