---
title: "22_analysis_18"
format: html
editor: visual
---

## Running Code

#### Loading libraries:

```{r}

library("dplyr")     
library("ggplot2")     
library("scales")
```

#### Read data:

```{r}

dat <- read.delim("../data/03_dat_aug.tsv")
```

#### Re-ordering

Ordering the different genotypes into groups from the augmented data:

```{r}

# Create ER/PR/HER2 combo groups
dat <- dat |>
  mutate(
    ER   = if_else(er_status  == "P",
                   "ER+", 
                   "ER-"),
    PR   = if_else(pr_status  == "P", 
                   "PR+", 
                   "PR-"),
    HER2 = if_else(her2_status == "P",
                   "HER2+",
                   "HER2-"),
    combo = paste(ER,
                  PR,
                  HER2,
                  sep = ", ")
  ) |>
  mutate(
    combo = factor(
      combo,
      levels = c(
        "ER+, PR-, HER2-",
        "ER+, PR+, HER2-",
        "ER+, PR-, HER2+",
        "ER+, PR+, HER2+",
        "ER-, PR+, HER2-",
        "ER-, PR+, HER2+",
        "ER-, PR-, HER2+",
        "ER-, PR-, HER2-"
      )
    )
  )

# check: how many rows total?
nrow(dat)

# check: how many patients in each genotype?
dat |> count(combo)
```

#### **Graph 5:**

##### **Prep: treatment among genotypes**

```{r}

# Clean treatment variable
dat_treat <- dat |>
  mutate(
    treatment = as.character(treatment),
    treatment = if_else(is.na(treatment), 
                        "NA",
                        treatment)
  )

# Summarise: count + proportions per genotype × treatment
treat_summary <- dat_treat |>
  group_by(combo,
           treatment) |>
  summarise(n = n(),
            .groups = "drop") |>
  group_by(combo) |>
  mutate(
    total = sum(n),
    pct   = n / total
  ) |>
  ungroup()

# Keep the same genotype order you already defined
treat_summary$combo <- factor(
  treat_summary$combo,
  levels = levels(dat$combo)
)
```

##### **Graph -** Treatment Distribution Across ER/PR/HER2 Genotypes

```{r}

treatment_colors <- c(
  "FAC"        = "#FF9999",
  "FACT"       = "#E69F00",
  "FACT+XRT/X" = "#B8860B",
  "FEC"        = "#66A61E",
  "FECT"       = "#1B9E77",
  "NA"         = "grey70",  
  "TFAC"       = "#00A6D6",
  "TFAC/HT"    = "#56B4E9",
  "TFEC"       = "#0072B2",
  "TH/FAC"     = "#7570B3",
  "TH/FEC"     = "#CC79A7",
  "Tonly"      = "#F0027F",
  "TXFAC"      = "#E7298A"
)

ggplot(treat_summary, aes(x = combo, y = pct, fill = treatment)) +
  
  # N labels above bars 
  geom_text(
    data = group_counts_treat,
    aes(
      x = combo,
      y = 1.07,                     
      label = paste0("N=", N_group)
    ),
    size = 3,
    inherit.aes = FALSE
  ) +
  
  # Stacked bars
  geom_col(position = "fill") +
  
  # Percent labels
  geom_text(
    aes(
      label = ifelse(pct > 0.02,
                     paste0(round(pct * 100), "%"),
                     "")
    ),
    position = position_fill(vjust = 0.5),
    size = 2.5,
    color = "black",
  ) +
  
  # Apply grey for NA + colors for other treatments
  scale_fill_manual(values = treatment_colors,
                    drop = FALSE) +

  scale_x_discrete(drop = FALSE) +
  scale_y_continuous(labels = percent_format()) +
  labs(
    title = "Treatment Distribution Across ER/PR/HER2 Genotypes",
    x = "ER/PR/HER2 Genotype",
    y = "Proportion of Patients",
    fill = "Treatment"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 20,
                               hjust = 1,
                               size = 7),
    plot.title  = element_text(size = 10),
    axis.title  = element_text(size = 9),

    legend.position = "bottom",
    legend.title = element_text(size = 7),
    legend.text  = element_text(size = 6),

    legend.box       = "horizontal",
    legend.box.spacing = unit(0.2,
                              "lines"),
    legend.key.size  = unit(3.5, "mm"),
    legend.key.width = unit(3.5, "mm"),
    legend.spacing.x = unit(0.5, "mm"),
    legend.spacing.y = unit(0.2, "mm")
  ) +
  
  guides(
    fill = guide_legend(
      ncol = 7,
      byrow = TRUE,
      title.position = "top"
    )
  ) +
  coord_cartesian(ylim = c(0,
                           1.15))
```

##### Conclusion:

Generally, more different treatments were tried in the last two groups, especially for the second last group, which also had the best response. The second group from the left (ER+, PR+, HER2-) also had more treatments but showed a bad response. So in conclusion, (1) more personalized treatment options give better response, (2) and a general broad TFAC and TFEC is not working for estrogen-positive patients.

Aggressiveness, tumor size, and nodes do not necessarily mean a bad outcome. With personalized treatments, it can mean good results for the patient even if the stage or grade of the tumor seems to have a theoretically bad prognosis.
